---
title: "Earthquake Analysis of Türkiye (2020-2026)"
author: "Özge Torcu Kılıç"
format: html
output-file: index.html
---

Türkiye is located on major fault lines and is highly prone to seismic activity. This project analyzes earthquakes with magnitude greater than 3 between 2020 and 2026 using data obtained from AFAD. The aim is to identify temporal trends and spatial patterns in earthquake occurrences.

# 1. Project Overview and Scope

In this project, I analyze earthquake data from Türkiye obtained from AFAD (Disaster and Emergency Management Presidency) for the years 2020-2026.

My aim is to show, using analytical methods, how seismic activity has changed, particularly after the 2023 Kahramanmaraş earthquakes, and in which regions it has clustered.

# 2. Data

The dataset was provided by AFAD and consists of a total of 9 columns.

The key columns are: `Magnitude`, `Latitude-Longitude`, and `Date`.

## 2.1 Data Source

<https://deprem.afad.gov.tr/event-catalog>

## 2.2 General Information About Data

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: asis

library(readxl)

deprem_data <- read_excel("data/deprem_verisi.xlsx")
deprem_data$Magnitude <- as.numeric(deprem_data$Magnitude)

cat(
  "This dataset contains earthquake records in Türkiye between 2020 and 2026, sourced from AFAD. It includes ",
  nrow(deprem_data),
  " observations and ",
  ncol(deprem_data),
  " variables. The magnitude values range between ",
  min(deprem_data$Magnitude, na.rm = TRUE),
  " and ",
  max(deprem_data$Magnitude, na.rm = TRUE),
  "."
)
```


## 2.3 Reason of Choice

Türkiye is a country deeply affected by seismic activities due to its geographical location on major fault lines. The 2023 Kahramanmaraş earthquakes highlighted the importance of data-driven disaster management. I chose this dataset to analyze earthquake patterns using analytical tools and visualizations.

## 2.4 Preprocessing

The raw data required several preparation steps before analysis:

- The dataset was imported from an Excel file using the `readxl` library.
- The `Date` column was converted into `POSIXct` format to enable time-based analysis.
- The `Magnitude` column was converted into numeric format.
- The cleaned dataset was saved as an `.RData` file for reproducibility.

```{r}
#| label: data-preparation
#| echo: false
#| message: false
#| warning: false

library(readxl)
library(dplyr)

deprem_data <- read_excel("data/deprem_verisi.xlsx")

deprem_data$Date <- as.POSIXct(
  deprem_data$Date,
  format = "%d/%m/%Y %H:%M:%S"
)

deprem_data$Magnitude <- as.numeric(deprem_data$Magnitude)

save(deprem_data, file = "deprem_verisi.RData")
```


# 3. Analysis

## 3.1 Exploratory Data Analysis

In this section, I perform an initial exploration of the earthquake data by visualizing the distribution of earthquake magnitudes.

```{r}
#| echo: false

library(ggplot2)

ggplot(deprem_data, aes(x = Magnitude)) +
  geom_histogram(binwidth = 0.5, fill = "red", color = "white") +
  labs(
    title = "Magnitude Distribution of Earthquakes in Türkiye (2020-2026)",
    x = "Magnitude",
    y = "Count"
  )
```

## 3.2 Trend Analysis

This section examines how earthquake frequency has changed over time.

```{r}
#| echo: false

library(dplyr)
library(ggplot2)

yearly_counts <- deprem_data %>%
  mutate(Year = format(Date, "%Y")) %>%
  count(Year)

ggplot(yearly_counts, aes(x = Year, y = n, group = 1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Number of Earthquakes by Year",
    x = "Year",
    y = "Count"
  )
```

## 3.3 Model Fitting

For the advanced analysis part, I may explore clustering methods such as K-means based on latitude and longitude to identify geographic hotspots of seismic activity. I may also examine simple time-based aggregations to better understand changes in earthquake frequency over time.

## 3.4 Results

Initial findings suggest that seismic activity in Türkiye is concentrated along major fault lines and that earthquake frequency varies across years.

# 4. Results and Key Takeaways

The frequency of earthquakes in Türkiye shows a noticeable increase after 2023.

Seismic activities are not uniformly distributed but are heavily clustered along the East and North Anatolian fault lines.

These patterns can provide useful insights for regional risk assessment and disaster preparedness.

------------------------------------------------------------------------

### Data Download

Click the link below to download the processed dataset for reproducibility:

[Download deprem_verisi.RData](deprem_verisi.RData)
