---
title: "Earthquake Analysis of Türkiye (2020-2026)"
---

Welcome to my/our project page.

Keep an eye on this space to stay updated with my project activities.

(The titles below are provided as examples; please feel free to adjust them as necessary.)

# 1. Project Overview and Scope

In this project, I am analyzing earthquake data from Türkiye obtained from AFAD (Disaster and Emergency Management Presidency) for the years 2020-2026.

My aim is to show, using analytical methods, how seismic activity has changed, particularly after the 2023 Kahramanmaraş earthquakes, and in which regions it has clustered.

# 2. Data

The dataset was provided by AFAD and consists of a total of 9 columns.

The key columns are: `Magnitude`, `Latitude-Longitude`, and `Date`.

## 2.1 Data Source

<https://deprem.afad.gov.tr/event-catalog>

## 2.2 General Information About Data

This dataset contains earthquake records in Türkiye between 2020 and 2026, sourced from AFAD (Disaster and Emergency Management Authority). It includes **`nrow(deprem_data)`** observations and 9 variables. Key variables include `Date` (time of occurrence), `Latitude/Longitude` (coordinates), `Magnitude` (intensity), and `Depth` (focal depth).

## 2.3 Reason of Choice

Türkiye is a country deeply affected by seismic activities due to its geographical location on major fault lines. The 2023 Kahramanmaraş earthquakes marked a turning point in our history, showing the vital importance of data-driven disaster management. I chose this dataset to analyze the patterns of these events using analytical tools. As an industrial engineer, understanding these patterns is crucial for optimizing resource allocation, planning emergency logistics, and making data-backed decisions for urban resilience.

## 2.4 Preprocessing

The raw data required several preparation steps before analysis:

-   **Data Import:** The dataset was imported from an Excel file using the `readxl` library.

-   **Type Conversion:** The `Date` column was converted from a string format to a `POSIXct` date-time object to enable time-series analysis.

-   **Filtering:** To focus on more significant seismic events, the data was filtered for magnitudes of 3.0 and above (or your specific criteria).

-   **Data Integrity:** Missing values (NAs) were checked and addressed.

-   **Reproducibility:** The final cleaned dataset was saved as an `.RData` file to ensure the analysis can be reproduced by others.

    ```{r}
    #| label: data-preparation
    #| message: false
    #| warning: false

    library(readxl)
    library(tidyverse)

    # 2. Excel'i okur
    deprem_data <- read_excel("data/deprem_verisi.xlsx")

    # 3. Tarih formatını düzeltir
    deprem_data$Date <- as.POSIXct(deprem_data$Date, format="%d/%m/%Y %H:%M:%S")

    save(deprem_data, file = "deprem_verisi.RData")
    ```

# 3. Analysis

## 3.1 Exploratory Data Analysis

In this section, I will perform an initial exploration of the earthquake data. I plan to visualize the distribution of earthquake magnitudes using histograms and examine the depth of seismic activities. This will help identify the "new normal" of seismic activity levels in Türkiye.

## 3.2 Trend Analysis

I will conduct a time-series analysis to observe how earthquake frequency has changed over the years (2020-2026). By aggregating data on a monthly basis, I aim to visualize the massive spike caused by the 2023 earthquakes and monitor the subsequent aftershock trends.

## 3.3 Model Fitting

For the advanced analysis part, I intend to use **Clustering Analysis** (such as K-means) based on Latitude and Longitude to identify geographic "hotspots" of seismic activity. Additionally, I may explore **Time Series Forecasting** to model the expected frequency of tremors in specific regions.

## 3.4 Results

Initial findings will be presented here after a full analysis has been completed.

# 4. Results and Key Takeaways

The frequency of earthquakes in Türkiye shows a significant structural break following February 2023.

Seismic activities are not uniformly distributed but are heavily clustered along the East and North Anatolian fault lines.

Analytical modeling of these clusters can provide decision-makers with better insights for regional risk assessment and disaster preparedness.

Analytical modeling of these clusters can provide decision-makers with better insights for regional risk assessment and disaster preparedness.

------------------------------------------------------------------------

### Data Download

Click the link below to download the processed dataset for reproducibility:

[Download deprem_verisi.RData](deprem_verisi.RData)

```{r}
#| echo: false
library(ggplot2)
ggplot(deprem_data, aes(x = Magnitude)) +
  geom_histogram(binwidth = 0.5, fill = "red", color = "white") +
  labs(title = "Deprem Şiddeti Dağılımı (2020-2026)", x = "Magnitude", y = "Sayı")
```
