---
title: "Assignment 1"
---

# Assignment 1

```{r}
```

My first assignment has two parts.

## (a)

Video: Conversations on Data Science and Industrial Engineering - Baykal Hafızoğlu & Erdi Daşdemir

Summary: The participants' academic backgrounds and how they transitioned from Industrial Engineering to Data Science were discussed. The close relationship between Industrial Engineering and Data Science was explained. It was emphasized that simply writing excellent code or creating complex mathematical models is not enough. The importance of transforming data into value was highlighted. It was stressed that data should be used to plan for the future and integrate it into processes, and that data science is a helpful tool in decision-making. It was stated that skills should be developed in line with understanding and application. Industry applications and challenges were discussed.

## (b)

Question 1: What is the biggest advantage of industrial engineers in data science?

It is their ability to think "systemically" and to manage data not only by analyzing it, but also as a decision-making and optimization process.

Question 2: What is required for a project to be considered "successful"?

A)  Using the newest and most expensive software.

B)  Having a very large amount of data.

C)  The model producing a useful result and solving a problem.

D)  The graphs looking very nice.

```{r}
#| label: task-3-c
#| echo: true

my_first <- "Özge"
my_birth_year <- 1996

# k sayısı hesaplama
k <- (nchar(my_first) + my_birth_year) %% 15 + 8
cat("calculated my k :", k, "\n")

# function
compute_stats <- function(x) {
  list(
    mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    variance = var(x, na.rm = TRUE),
    iqr = IQR(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE)
  )
}

# 1. using sapply
cat("\n--- sapply Sonuçları ---\n")
sapply(mtcars, compute_stats)

# 2. using apply
cat("\n--- apply Sonuçları ---\n")
apply_results <- apply(as.matrix(mtcars), 2, compute_stats)
print(apply_results)

# for döngüsü ile mtcars analizi 
cat("\n--- For Döngüsü Sonuçları ---\n")
for (i in 1:ncol(mtcars)) {
  cat(colnames(mtcars)[i], ":\n")
  print(compute_stats(mtcars[[i]]))
}
```

```{r}
#| label: task-3-d
#| echo: true
library(dslabs)
data("polls_us_election_2016")

# k sayısına göre (k=13, TEK) son 13 satırı gösterir
cat("\nOrijinal Verinin Son k Satırı:\n")
if (k %% 2 == 0) {
  print(head(polls_us_election_2016, k))
} else {
  print(tail(polls_us_election_2016, k))
}

# sum NA 
cat("\nToplam NA sayısı:", sum(is.na(polls_us_election_2016)), "\n")

# NA Değerlerini Doldurma
new_data <- polls_us_election_2016

for(i in 1:ncol(new_data)) {
  if(is.numeric(new_data[[i]])) {
    new_data[[i]][is.na(new_data[[i]])] <- my_birth_year + k
  } else {
    new_data[[i]] <- as.character(new_data[[i]])
    new_data[[i]][is.na(new_data[[i]])] <- paste0(my_first, "_", k)
  }
}

# Yeni veri setinin kontrolü
cat("\nDoldurma Sonrası Yeni Verinin Son k Satırı:\n")
if (k %% 2 == 0) {
  print(head(new_data, k))
} else {
  print(tail(new_data, k))
}

cat("\nDoldurma sonrası boş veri var mı?:", anyNA(new_data))
```

# 

> **AI Disclosure:** I utilized AI tools (Gemini/ChatGPT) to assist in structuring the R code and translating the assignment requirements.
>
> **Prompt used:** "Could you write R code for my EMU660 assignment that calculates the number k, analyzes mtcars data, and populates the NA values?"
